### 2009-10-03 paper/paper.tex
Up!

\documentclass{letter}[11pt]

\begin{document}
\begin{letter}{

Test stuff
\end{document}

### 2009-10-18 paper/paper.tex
Stuff

@@
-\documentclass{letter}[11pt]
+\title{ A Traveling Salesman Solution For The Capitals of All African Nations }
+\author{
+        Brian Gianforcaro \\
+                Department of Computer Science\\
+        Rochester Institute of Technology\\
+}
+
+\date{\today}
+
+\documentclass[12pt]{article}
+
+\usepackage{graphicx}
+\usepackage{multicol}
+\usepackage{mdwlist}
 
 \begin{document}
-\begin{letter}{
+\maketitle
 
-Test stuff
+\begin{abstract}
+A Traveling Salesman Problem is the task of finding 
+the shortest round trip path a traveling salesperson can take 
+to visit each vertex of a given graph.  They are usually 
+implemented using a genetic algorithm. Our salesperson
+happens to be traveling to the capitals of every country in Africa that is a recognized member of the United Nations.
+\end{abstract}
+
+\section{The Problem}
+
+\begin{figure}
+\begin{center}
+\setlength\fboxsep{1.00pt}
+\setlength\fboxrule{1.00pt}
+\fbox{\includegraphics[scale=0.80]{points.png}}
+\caption{Capitals of African Nations}
+\end{center}
+\end{figure}
+
+\begin{multicols}{3}
+\tiny\begin{enumerate*}
+\item Algeria - Algiers
+\item Angola - Luanda
+\item Benin - Porto-Novo
+\item Botswana - Gaborone
+\item Burkina Faso - Ouagadougou
+\item Burundi - Bujumbura
+\item Cameroon - Yaounde
+\item Cape Verde - Praia
+\item Central African Republic - Bangui
+\item Chad - N'Djamena
+\item Comoros - Moroni
+\item Congo, Republic of the - Brazzaville
+\item Congo, Democratic Republic of the - Kinshasa
+\item Cote d'Ivoire - Yamoussoukro
+\item Djibouti - Djibouti
+\item Egypt - Cairo
+\item Equatorial Guinea - Malabo
+\item Eritrea - Asmara
+\item Ethiopia - Addis Ababa
+\item Gabon - Libreville
+\item The Gambia - Banjul
+\item Ghana - Accra
+\item Guinea - Conakry
+\item Guinea-Bissau - Bissau
+\item Kenya - Nairobi
+\item Lesotho - Maseru
+\item Liberia - Monrovia
+\item Libya - Tripoli
+\item Madagascar - Antananarivo
+\item Malawi - Lilongwe
+\item Mali - Bamako
+\item Mauritania - Nouakchott
+\item Mauritius - Port Louis
+\item Morocco - Rabat
+\item Mozambique - Maputo
+\item Namibia - Windhoek
+\item Niger - Niamey
+\item Nigeria - Abuja
+\item Rwanda - Kigali
+\item Senegal - Dakar
+\item Seychelles - Victoria
+\item Sierra Leone - Freetown
+\item Somalia - Mogadishu
+\item South Africa - Pretoria 
+\item Sudan - Khartoum
+\item Swaziland - Mbabane
+\item Tanzania - Dar es Salaam
+\item Togo - Lome
+\item Tunisia - Tunis
+\item Uganda - Kampala
+\item Zambia - Lusaka
+\item Zimbabwe - Harare
+\end{enumerate*}
+\end{multicols}
+
+\section{Overview}
+The remainder of this article is organized as follows.
+Section~ gives account of previous work.
+Our new and exciting results are described in Section~.
+Finally, Section gives the conclusions.
+
+\section{Programs}
+A much longer  example was written by Gil
+
+\section{Solution}
+
+\begin{figure}
+\begin{center}
+\setlength\fboxsep{1.00pt}
+\setlength\fboxrule{1.00pt}
+\fbox{\includegraphics[scale=0.65]{path.png}}
+\caption{Final Path Through Africa}
+\end{center}
+\end{figure}
+
+In this section we describe the results.
+
+\section{Runtime}
+We worked hard, and achieved very little.
+
+\section{Analysis}
+
 \end{document}

### 2009-10-18 code/hillclimb.py

@@ -1,4 +1,6 @@
 import logging
+import os
+from urllib import urlretrieve 
 
 def hillclimb(init_function,move_operator,objective_function,max_evaluations):
     '''
@@ -33,7 +35,7 @@ def hillclimb(init_function,move_operator,objective_function,max_evaluations):
     logging.info('hillclimb finished: num_evaluations=%d, best_score=%f',num_evaluations,best_score)
     return (num_evaluations,best_score,best)
 
-def hillclimb_and_restart(init_function,move_operator,objective_function,max_evaluations):
+def hillclimb_and_restart(init_function,move_operator,objective_function,max_evaluations,coords):
     '''
     repeatedly hillclimb until max_evaluations is reached
     '''
@@ -52,5 +54,26 @@ def hillclimb_and_restart(init_function,move_operator,objective_function,max_eva
             best_score=score
             best=found
             logging.critical("New best: "+ str(best_score)+ " remaining: %d/%d",remaining_evaluations,max_evaluations)
+            getimg(coords,best)
+            os.popen("open path.png")
+
         
     return (num_evaluations,best_score,best)
+
+def getimg(coords, best):
+    url = "http://maps.google.com/maps/api/staticmap?path=color:orange|weight:4"
+
+    for i in best:
+        url = url + "|" + str(coords[i][0]) + "," + str(coords[i][1])
+
+    url = url + "&size=1024x1024&sensor=false&key=ABQIAAAAzr2EBOXUKnm_jVnk0OJI7xSsTL4WIgxhMZ0ZK_kHjwHeQuOD4xQJpBVbSrqNn69S6DOTv203MQ5ufA"
+    urlretrieve(url,"path.png")
+
+    url = "http://maps.google.com/maps/api/staticmap?markers=color:blue|"
+    for i in best:
+            url = url + "|" + str(coords[i][0]) + "," + str(coords[i][1])
+
+    url = url + "&size=1024x1024&sensor=false&key=ABQIAAAAzr2EBOXUKnm_jVnk0OJI7xSsTL4WIgxhMZ0ZK_kHjwHeQuOD4xQJpBVbSrqNn69S6DOTv203MQ5ufA"
+    urlretrieve(url,"points.png")
+
+

### 2009-10-18 code/tsp.py

@@ -110,9 +110,9 @@ def init_random_tour(tour_length):
    random.shuffle(tour)
    return tour
 
-def run_hillclimb(init_function,move_operator,objective_function,max_iterations):
+def run_hillclimb(init_function,move_operator,objective_function,max_iterations,coords):
     from hillclimb import hillclimb_and_restart
-    iterations,score,best=hillclimb_and_restart(init_function,move_operator,objective_function,max_iterations)
+    iterations,score,best=hillclimb_and_restart(init_function,move_operator,objective_function,max_iterations,coords)
     return iterations,score,best
 
 def usage():
@@ -190,7 +190,7 @@ def main():
     
     logging.info('using move_operator: %s'%move_operator)
     
-    iterations,score,best=run_hillclimb(init_function,move_operator,objective_function,max_iterations)
+    iterations,score,best=run_hillclimb(init_function,move_operator,objective_function,max_iterations,coords)
     # output results
     # print coords
     getimg(coords,best)

### 2009-10-31 paper/paper.tex
added spelling check
stuff

@@
 \usepackage{mdwlist}
+\usepackage{amssymb,amsmath}
+\usepackage{cite}
+\usepackage[bookmarks=false,colorlinks=true,linkcolor={blue},pdfstartview={XYZ null null 1.22}]{hyperref} 
+\usepackage{url}
 
@@
 \section{Programs}
-A much longer  example was written by Gil
+
+The TSP was solved using the Python 2.6 programming language.
+
+I leveraged software written by John Montgomery \cite{cite_key2}
+
+The results were then visualized using Google maps mapping API.
+
+Two methods were
 
 \section{Solution}
+
+\begin{align}
+  yDis = ( lat2 - lat1 ) * Nautical Miles Per Latitude\\
+  xDis = ( \cos( lat1 - \frac{\pi}{180} ) + \cos( lat2 - \frac{\pi}{180} ) ) * ( lon2 - lon1 ) * \frac{Nautical Miles Per Longitude}{2}\\
+  tDistance = \sqrt{ yDis^{2} + xDis^{2} } * Miles per Nautical Miles\\
+\end{align}
 
@@
 \section{Analysis}
-
+\begin{thebibliography}{99}
+\bibitem{cite_key1} Montgomery, John {\it Tackling The Travelling Salesman Problem}{\href{http://www.psychicorigami.com/category/tsp/}{ http://www.psychicorigami.com/category/tsp/} }, {\bf 2007}
+\bibitem{cite_key2} Mead, C. A.; Truhlar, D. G. {\it J. Chem. Phys.} {\bf 1983}, {\it 78}, 6344.
+\end{thebibliography}
 \end{document}

### 2009-11-01 fastpath/fastpath.py
Stuff

@@ -1,0 +1,113 @@
+#!/usr/bin/python
+
+import random
+import sys
+import getopt
+from math import sqrt
+import math
+
+nauticalMilePerLat = 60.00721
+nauticalMilePerLongitude = 60.10793
+rad = math.pi / 180.0
+milesPerNauticalMile = 1.15078
+
+def read_coords(coord_file):
+    '''
+    read the coordinates from file and return the distance matrix.
+    coords should be stored as comma separated floats, one x,y pair per line.
+    '''
+    coords=[]
+    for line in coord_file:
+        x,y=line.strip().split(',')
+        coords.append((float(x),float(y)))
+    return coords
+
+def order(best):
+    file = open('cities','r')
+    cities = []
+    for city in file:
+      cities.append(city)
+
+    file.close()
+
+    for i in best:
+      print cities[i]
+
+def calcDistance(lat1, lon1, lat2, lon2):
+  """
+  Caclulate distance between two lat lons in NM
+  """
+  yDistance = (lat2 - lat1) * nauticalMilePerLat
+  xDistance = (math.cos(lat1 * rad) + math.cos(lat2 * rad)) * (lon2 - lon1) * (nauticalMilePerLongitude / 2)
+  distance = math.sqrt( yDistance**2 + xDistance**2 )
+  return distance * milesPerNauticalMile
+
+def calculatePathLen(path, arr):
+  first = True
+  lastpoint = ()
+  length = 0
+  for point in path:
+    if first:
+      lastpoint = point
+      first = False
+    else:
+      length += calcDistance(arr[lastpoint][0],arr[lastpoint][1],arr[point][0],arr[point][1])
+      lastpoint = point
+
+  return length
+
+def usage():
+    print "usage: python %s [-v] -n <max iterations> <city file>" % sys.argv[0]
+
+def main():
+    try:
+        options, args = getopt.getopt(sys.argv[1:], "ho:vm:n:a:", ["cooling="])
+    except getopt.GetoptError:
+        usage()
+        sys.exit(2)
+    max_iterations=None
+    verbose=None
+    for option,arg in options:
+        if option == '-v':
+            verbose=True
+        elif option == '-h':
+            usage()
+            sys.exit()
+        elif option == '-n':
+            max_iterations=int(arg)
+
+    if max_iterations is None:
+        usage();
+        sys.exit(2)
+
+    if out_file_name and not out_file_name.endswith(".png"):
+        usage()
+        print "output image file name must end in .png"
+        sys.exit(1)
+
[…]

### 2009-11-01 fastpath/test.py

@@ -1,0 +1,157 @@
+import random
+import sys
+
+INFINITE=1000000000
+
+class Node:
+  def __init__(self, pos, open=True):
+    assert(len(pos)==2)
+    self.pos = pos        #where this node is located
+    self.neighbours = []    #a list of all neighbour nodes
+    self.open = open      #True if this node is traversable
+    self.distance = 0     #distance to the startnode
+    self.visited = False    #True if this node has already been visited this run
+    self.visited_from = None  # the tile we came from when calculating distance, not needed as were not interested in the path 
+
+  def __repr__(self):
+    s = "pos: "+str(self.pos)
+    s += "\nopen: "+str(self.open)
+    s += "\ndistance: "+str(self.distance)
+    s += "\nvisited: "+str(self.visited)
+    return s
+
+  def add_neighbour(self, n):
+    if not n in self.neighbours and n is not self and self.open and n.open:
+      self.neighbours.append(n)
+
+class Graph:
+  def __init__(self):
+    self.nodes = []
+    self.longest_distance = (0,None,None) #three element tuple; distance, node1, node2
+
+  def add_node(self, x, y):
+    self.nodes.append(Node((x,y)))
+
+  def find_node(self, pos):
+    for i in self.nodes:
+      if i.pos==pos:
+        return i
+    return None
+
+  def reset(self, startpos):
+    for i in self.nodes:
+      i.distance = INFINITE
+      i.visited = False
+
+  def calc_neighbours(self):
+    print "calculating neigbours"
+    for i in self.nodes:
+      for p in self.nodes:
+          i.add_neighbour(p)
+
+  def dijkstra(self, startpos):
+    print "running dijkstra from",startpos
+    self.reset(startpos)
+    current = self.find_node(startpos)
+    current.distance = 0
+
+    #run as long as we have nodes to visit
+    while current:
+      for i in current.neighbours:
+        if not i.visited:
+          dist = current.distance+1
+          if dist < i.distance:
+            i.distance = dist
+            i.visited_from = current
+      current.visited = True
+      current = self.find_next_unvisited()
+
+    #update longest distance
+    longest = None
+    for i in self.nodes:
+      if longest==None or i.distance > longest.distance:
+        longest = i
+    if longest.distance > self.longest_distance[0]:
[…]

### 2009-11-01 hillclimb/tsp.py

@@ -64,8 +64,8 @@ def tour_length(matrix,tour):
     '''total up the total length of the tour based on the distance matrix'''
     total=0
     num_cities=len(tour)
-    for i in range(num_cities):
-        j=(i+1)%num_cities
+    for i in range(num_cities-1):
+        j=i+1
         city_i=tour[i]
         city_j=tour[j]
         total+=matrix[city_i,city_j]

### 2009-11-01 test/hillclimb.py

@@ -1,6 +1,6 @@
 import logging
 import os
-from urllib import urlretrieve 
+from urllib import urlretrieve
 
 def hillclimb(init_function,move_operator,objective_function,max_evaluations):
     '''
@@ -54,8 +54,25 @@ def hillclimb_and_restart(init_function,move_operator,objective_function,max_eva
             best_score=score
             best=found
             logging.critical("New best: "+ str(best_score)+ " remaining: %d/%d",remaining_evaluations,max_evaluations)
-            #getimg(coords,best)
-            #os.popen("open path.png")
+            getimg(coords,best)
+            os.popen("open path.png")
 
-        
     return (num_evaluations,best_score,best)
+
+def getimg(coords, best):
+    url = "http://maps.google.com/maps/api/staticmap?path=color:orange|weight:4"
+
+    for i in best:
+        url = url + "|" + str(coords[i][0]) + "," + str(coords[i][1])
+
+    url = url + "&size=1024x1024&sensor=false&key=ABQIAAAAzr2EBOXUKnm_jVnk0OJI7xSsTL4WIgxhMZ0ZK_kHjwHeQuOD4xQJpBVbSrqNn69S6DOTv203MQ5ufA"
+    urlretrieve(url,"path.png")
+
+    url = "http://maps.google.com/maps/api/staticmap?markers=color:blue|"
+    for i in best:
+            url = url + "|" + str(coords[i][0]) + "," + str(coords[i][1])
+
+    url = url + "&size=1024x1024&sensor=false&key=ABQIAAAAzr2EBOXUKnm_jVnk0OJI7xSsTL4WIgxhMZ0ZK_kHjwHeQuOD4xQJpBVbSrqNn69S6DOTv203MQ5ufA"
+    urlretrieve(url,"points.png")
+
+

### 2009-11-01 test/sa.py

@@ -25,7 +25,7 @@ class ObjectiveFunction:
             self.best_score=score
             self.best=solution
             logging.info('new best score: %f',self.best_score)
-            if score > (-341):
+            if score > (-325):
               getimg(self.coords,self.best)
               os.popen("eog path.png")
 

### 2009-11-01 test/tsp.py

@@ -84,15 +84,23 @@ def order(best):
       print cities[i]
 
 
-def tour_length(matrix,tour):
+def tour_length(matrix, tour, tsp):
     '''total up the total length of the tour based on the distance matrix'''
     total=0
     num_cities=len(tour)
-    for i in range(num_cities):
-        j=(i+1)%num_cities
-        city_i=tour[i]
-        city_j=tour[j]
-        total+=matrix[city_i,city_j]
+    if tsp:
+      for i in range(num_cities):
+          j=(i+1)%num_cities
+          city_i=tour[i]
+          city_j=tour[j]
+          total+=matrix[city_i,city_j]
+    else:
+      for i in range(num_cities-1):
+          j=i+1
+          city_i=tour[i]
+          city_j=tour[j]
+          total+=matrix[city_i,city_j]
+
     return total
 
 def calcDistance(lat1, lon1, lat2, lon2):
@@ -138,11 +146,11 @@ def run_anneal(init_function,move_operator,objective_function,max_iterations,sta
     return iterations,score,best
 
 def usage():
-    print "usage: python %s [-o <output image file>] [-v] [-m reversed_sections|swapped_cities] -n <max iterations> [-a hillclimb|anneal] [--cooling start_temp:alpha] <city file>" % sys.argv[0]
+    print "usage: python %s [-o <output image file>] [-v] [-m reversed_sections|swapped_cities]  -n <max iterations> [-a hillclimb|anneal] -t [--cooling start_temp:alpha] <city file>" % sys.argv[0]
 
 def main():
     try:
-        options, args = getopt.getopt(sys.argv[1:], "ho:vm:n:a:", ["cooling="])
+      options, args = getopt.getopt(sys.argv[1:], "ho:vm:n:a:t:", ["cooling="])
     except getopt.GetoptError:
         usage()
         sys.exit(2)
@@ -152,6 +160,8 @@ def main():
     move_operator=reversed_sections
     run_algorithm=run_hillclimb
     coords = []
+
+    tsp = False
     
     start_temp,alpha=None,None
     
@@ -170,6 +180,8 @@ def main():
                 move_operator=swapped_cities
             elif arg == 'reversed_sections':
                 move_operator=reversed_sections
+        elif option == '-t':
+            tsp = True
         elif option == '-a':
             if arg == 'hillclimb':
                 run_algorithm=run_hillclimb
@@ -211,13 +223,16 @@ def main():
     coords=read_coords(file(city_file))
[…]

### 2009-11-01 test/sa.py
stuff

@@ -25,9 +25,9 @@ class ObjectiveFunction:
             self.best_score=score
             self.best=solution
             logging.info('new best score: %f',self.best_score)
-            if score > (-325):
+            if score > (-21700):
               getimg(self.coords,self.best)
-              os.popen("eog path.png")
+              os.popen("open path.png")
 
         return score
 

### 2009-11-01 test/tsp.py

@@ -56,9 +56,7 @@ def cartesian_matrix(coords):
     matrix={}
     for i,(x1,y1) in enumerate(coords):
         for j,(x2,y2) in enumerate(coords):
-            dx,dy=x1-x2,y1-y2
-            dist=sqrt(dx*dx + dy*dy)
-            matrix[i,j]=dist
+            matrix[i,j]=calcDistance(x1,y1,x2,y2)
     return matrix
 
 def read_coords(coord_file):
@@ -88,18 +86,17 @@ def tour_length(matrix, tour, tsp):
     '''total up the total length of the tour based on the distance matrix'''
     total=0
     num_cities=len(tour)
-    if tsp:
-      for i in range(num_cities):
-          j=(i+1)%num_cities
-          city_i=tour[i]
-          city_j=tour[j]
-          total+=matrix[city_i,city_j]
-    else:
-      for i in range(num_cities-1):
-          j=i+1
-          city_i=tour[i]
-          city_j=tour[j]
-          total+=matrix[city_i,city_j]
+    max = 0
+    for i in range(num_cities):
+      j = (i+1)%num_cities
+      city_i=tour[i]
+      city_j=tour[j]
+      total+=matrix[city_i,city_j]
+      if matrix[city_i,city_j] > max:
+          max = matrix[city_i,city_j]
+
+    if not tsp:
+       total = total - max
 
     return total
 
@@ -112,17 +109,24 @@ def calcDistance(lat1, lon1, lat2, lon2):
   distance = math.sqrt( yDistance**2 + xDistance**2 )
   return distance * milesPerNauticalMile
 
-def calculatePathLen(path, arr):
+def calculatePathLen(path, arr, tsp):
   first = True
   lastpoint = ()
   length = 0
+  max = 0
   for point in path:
     if first:
       lastpoint = point
       first = False
     else:
-      length += calcDistance(arr[lastpoint][0],arr[lastpoint][1],arr[point][0],arr[point][1])
+      curr = calcDistance(arr[lastpoint][0],arr[lastpoint][1],arr[point][0],arr[point][1])
+      length += curr
       lastpoint = point
+      if curr > max:
+        max = curr
+
+  if not tsp:
+    length = length - max
 
   return length
 
@@ -150,7 +154,7 @@ def usage():
 
 def main():
     try:
-      options, args = getopt.getopt(sys.argv[1:], "ho:vm:n:a:t:", ["cooling="])
[…]

### 2009-11-01 paper/paper.tex
Stuff
stuff
paper

--- 2009-10-18 paper/paper.tex
@@
 \usepackage{mdwlist}
+\usepackage{amssymb,amsmath}
+\usepackage{cite}
+\usepackage[bookmarks=false,colorlinks=true,linkcolor={blue},pdfstartview={XYZ null null 1.22}]{hyperref}
+\usepackage{url}
+\usepackage[T1]{fontenc}
+%\usepackage{gfsartemisia-euler}
 
@@
 
-\begin{figure}
+\begin{figure}[h!]
   \begin{center}
@@
     \setlength\fboxrule{1.00pt}
-\fbox{\includegraphics[scale=0.80]{points.png}}
+    \fbox{\includegraphics[scale=0.60]{points.png}}
     \caption{Capitals of African Nations}
@@
 \section{Overview}
-The remainder of this article is organized as follows.
-Section~ gives account of previous work.
-Our new and exciting results are described in Section~.
-Finally, Section gives the conclusions.
+
+  The traveling salesman problem (TSP) is a thoroughly researched problem in theoreticaly computer-science. 
+TSP is classified as NP-Hard eg. as hard or harder to solve than a problem solvable in nondeterministic polynomial time.
+This means the worst case performance for a TSP algorithm will likely increase exponetially with the numbers of cities to traverse.
+
+The big-O complexity of a brute force TSP algorithms (check all vertices against all other vertices) is $O(n!)$.
 
 \section{Programs}
-A much longer  example was written by Gil
+
+\begin{enumerate}
+\item The Python 2.6 \cite{python} programming language and interpreter.
+\item Wikipedia's list of city latitude/longitudes \cite{wiki}.
+\item Software originally written by John Montgomery \cite{tsp} in 2007.General exercise in different types of TSP solving algorithms.
+\item The results were then visualized using Google maps static mapping API \cite{google}.
+
+\end{enumerate}
 
@@
 
-\begin{figure}
+%$yDis = ( lat2 - lat1 ) * \text{Nautical Miles Per Latitude}$\newline
+%$xDis = ( \cos( lat1 - \frac{\pi}{180} ) + \cos( lat2 - \frac{\pi}{180} ) ) * ( lon2 - lon1 ) * \frac{\text{Nautical Miles Per Longitude}}{2}$\newline
+%$tDistance = \sqrt{ yDis^{2} + xDis^{2} } * \text{Miles per Nautical Miles}$\newline
+
+\begin{figure}[!h]
   \begin{center}
@@
     \setlength\fboxrule{1.00pt}
-\fbox{\includegraphics[scale=0.65]{path.png}}
+    \fbox{\includegraphics[scale=0.50]{path.png}}
     \caption{Final Path Through Africa}
@@
 
-In this section we describe the results.
+Final Order:
+\begin{multicols}{3}
+\tiny\begin{enumerate*}
+\item Congo, Republic of the - Brazzaville
+\item Congo, Democratic Republic of the - Kinshasa
+\item Angola - Luanda
+\item Namibia - Windhoek
+\item Botswana - Gaborone
+\item South Africa - Pretoria
+\item Lesotho - Maseru
+\item Swaziland - Mbabane
+\item Mozambique - Maputo
+\item Zimbabwe - Harare
+\item Zambia - Lusaka
+\item Malawi - Lilongwe
+\item Burundi - Bujumbura
+\item Rwanda - Kigali
+\item Uganda - Kampala
+\item Kenya - Nairobi
+\item Tanzania - Dar es Salaam
+\item Comoros - Moroni
+\item Madagascar - Antananarivo
+\item Mauritius - Port Louis
+\item Seychelles - Victoria
+\item Somalia - Mogadishu
+\item Ethiopia - Addis Ababa
+\item Djibouti - Djibouti
+\item Eritrea - Asmara
+\item Sudan - Khartoum
+\item Egypt - Cairo
+\item Libya - Tripoli
+\item Tunisia - Tunis
+\item Algeria - Algiers
+\item Morocco - Rabat
+\item Mauritania - Nouakchott
+\item Cape Verde - Praia
+\item Senegal - Dakar
+\item The Gambia - Banjul
+\item Guinea-Bissau - Bissau
+\item Guinea - Conakry
+\item Sierra Leone - Freetown
+\item Liberia - Monrovia
+\item Cote d'Ivoire - Yamoussoukro
+\item Mali - Bamako
+\item Burkina Faso - Ouagadougou
+\item Niger - Niamey
+\item Ghana - Accra
+\item Togo - Lome
+\item Benin - Porto-Novo
+\item Nigeria - Abuja
+\item Chad - N'Djamena
+\item Central African Republic - Bangui
+\item Cameroon - Yaounde
+\item Equatorial Guinea - Malabo
+\item Gabon - Libreville
+\end{enumerate*}
+\end{multicols}
+
+Total Round Trip $\approx 22253.2035387$ miles
 
 \section{Runtime}
-We worked hard, and achieved very little.
+
+The two algorithms, simulated annealing and brute force random permutations where both run in numerious instances over the life of the project. 
+In combined total they were run upwards of fity times a peice.
+
+The version of the algorithm wich used simulated annealing on average found the optimal solution in 6 seconds. 
+While the bruteforce method often took upwards of
 
@@
 
+I believe the final optimal path's are within a reasonable uncertenty of the actual path.
+Given the big-O of the brute force algorithm $O(n!)$ or $O(80658175170943878571660636856403766975289505440883277824000000000000)$ a runtime complexity of this magnitude is obviously out of my leauge for finding exact values.
+At a maxiumum I was able to attempt 100,000,000 permutations of the path.
+
+Two area's of possible improvement I can pinpoint are:
+\begin{enumerate}
+\item The language itself. I chose the python programming language for my TSP implementation.
+The language is interpreted and it's runtime speed can at times be 20\% slower than an equivilant C program.
+\item The random.shuffle implementation. Given how python's random.shuffle API is implemented it talkes a set
+of values and shuffles them based on a random number between 0,1 generated by the operating system. However given the size of the set
+, 52 individual points, the likely hood that random.shuffle will generated all permutations without excessive doubles is increadibly unlikely.
+This is the cause of a lot of literally useless computation time, however because of the size of the data set, caching data sets is not plausible let alone if even possible.
+\end{enumerate}
+
+Re-writing the algorithm in C would be a great performance benifit.
+Also having a proper resources to cache routes or a effective random route generater I beleive a more accurate optimal route might be found.
+
+\begin{thebibliography}{99}
+\bibitem{python} {\href{ http://python.org }{ http://python.org }}
+\bibitem{wiki} {\href{ http://en.wikipedia.org/wiki/Latitude_and_longitude_of_cities }{ http://en.wikipedia.org/wiki/Latitude\_and\_longitude\_of\_cities }}
+\bibitem{tsp} Montgomery, John {\it Tackling The Travelling Salesman Problem }{\href{ http://www.psychicorigami.com/category/tsp/ }{ http://www.psychicorigami.com/category/tsp/ } }, {\bf 2007}
+\bibitem{google} {\href{ http://code.google.com/apis/maps/documentation/staticmaps/  }{ http://code.google.com/apis/maps/documentation/staticmaps/ }}
+
+\end{thebibliography}
 \end{document}

### 2009-11-02 paper/paper.tex
stuff
Paper and new solutions

--- 2009-10-18 paper/paper.tex
@@
 
-\begin{figure}
+  Attempt to find the shortest i.e. cheapest complete path around all African Nations so our sales person can travel the most efficient route possible.
+
+\begin{figure}[h!]
   \begin{center}
@@
 \section{Overview}
-The remainder of this article is organized as follows.
-Section~ gives account of previous work.
-Our new and exciting results are described in Section~.
-Finally, Section gives the conclusions.
 
+  The traveling salesman problem (TSP) is a thoroughly researched problem in theoretically computer-science. 
+TSP is classified as NP-Hard e.g. it is as hard or harder to solve than a problem solvable in nondeterministic polynomial time.
+This means the worst case performance for a TSP algorithm will likely increase exponentially with the numbers of cities to traverse.
+The big-O complexity of a brute force TSP algorithms (check all vertices against all other vertices) is $O(n!)$.
+
+  I attempted to solve the TSP with both a brute force method initially and eventually a simulated annealing algorithm. 
+Both implementation's rely on the pythong programming language and it's random number implementation to generate new routes.
+The brute force algorithm simply attempts all combinations, giving priority to each new best performing route.
+This algorithm performed well, yielding great result's, but I felt it would be beneficial to try another approach. 
+Simulated annealing is often cited as a relatively easy way to improve TSP performance.
+
+\begin{quotation}\bf``\normalfont Simulated annealing starts with the cities connected in a random order, and then considers making random changes in that order.
+If changing the order of cities leads to a shorter path, we accept that change.
+If the modification yields a longer path, we give ourselves a certain probability of accepting the modification -- less likely the larger the proposed increase in path length.
+We then gradually reduce this probability over time, in order to rule out shorter and shorter path increases -- thereby converging toward a path length close to the absolute minimum.\bf''\normalfont\cite{sa} \end{quotation}
 \section{Programs}
-A much longer  example was written by Gil
+
+\begin{enumerate}
+\item The Python 2.6 \cite{python} programming language and interpreter.
+\item Wikipedia provides a convenient list of latitude/longitude pairs for all major cities in the world. This was cross referenced with
+the list of African nations and capitals recognized by the United Nations.\cite{wiki}
+\item Software originally written by John Montgomery \cite{tsp} in 2007.
+It was a general exercise in implementing different types of TSP algorithms.
+At this point it's been heavily modified to meet my needs for this project.
+\item The results were then visualized using Google maps static mapping API. \cite{google}
+\item A general algorithm for computing the distance between two latitude/longitude points.\cite{dis}
+\end{enumerate}
 
@@
     \setlength\fboxrule{1.00pt}
-\fbox{\includegraphics[scale=0.65]{path.png}}
-\caption{Final Path Through Africa}
+    \fbox{\includegraphics[scale=0.50]{path.png}}
+    \caption{Final Round Trip Path Through Africa}
   \end{center}
@@
 
-In this section we describe the results.
+Final Round Trip Order:
+\begin{multicols}{3}
+\tiny\begin{enumerate*}
+\item Congo, Republic of the - Brazzaville
+\item Congo, Democratic Republic of the - Kinshasa
+\item Angola - Luanda
+\item Namibia - Windhoek
+\item Botswana - Gaborone
+\item South Africa - Pretoria
+\item Lesotho - Maseru
+\item Swaziland - Mbabane
+\item Mozambique - Maputo
+\item Zimbabwe - Harare
+\item Zambia - Lusaka
+\item Malawi - Lilongwe
+\item Burundi - Bujumbura
+\item Rwanda - Kigali
+\item Uganda - Kampala
+\item Kenya - Nairobi
+\item Tanzania - Dar es Salaam
+\item Comoros - Moroni
+\item Madagascar - Antananarivo
+\item Mauritius - Port Louis
+\item Seychelles - Victoria
+\item Somalia - Mogadishu
+\item Ethiopia - Addis Ababa
+\item Djibouti - Djibouti
+\item Eritrea - Asmara
+\item Sudan - Khartoum
+\item Egypt - Cairo
+\item Libya - Tripoli
+\item Tunisia - Tunis
+\item Algeria - Algiers
+\item Morocco - Rabat
+\item Mauritania - Nouakchott
+\item Cape Verde - Praia
+\item Senegal - Dakar
+\item The Gambia - Banjul
+\item Guinea-Bissau - Bissau
+\item Guinea - Conakry
+\item Sierra Leone - Freetown
+\item Liberia - Monrovia
+\item Cote d'Ivoire - Yamoussoukro
+\item Mali - Bamako
+\item Burkina Faso - Ouagadougou
+\item Niger - Niamey
+\item Ghana - Accra
+\item Togo - Lome
+\item Benin - Porto-Novo
+\item Nigeria - Abuja
+\item Chad - N'Djamena
+\item Central African Republic - Bangui
+\item Cameroon - Yaounde
+\item Equatorial Guinea - Malabo
+\item Gabon - Libreville
+\end{enumerate*}
+\end{multicols}
+
+Total Length Of Round Trip Route $\approx 35,813 km$ \newline
+
+Route Without Round Trip Stipulation:
+\begin{multicols}{3}
+\tiny\begin{enumerate*}
+\item Benin - Porto-Novo
+\item Togo - Lome
+\item Ghana - Accra
+\item Niger - Niamey
+\item Burkina Faso - Ouagadougou
+\item Mali - Bamako
+\item Cote d'Ivoire - Yamoussoukro
+\item Liberia - Monrovia
+\item Sierra Leone - Freetown
+\item Guinea - Conakry
+\item Guinea-Bissau - Bissau
+\item The Gambia - Banjul
+\item Senegal - Dakar
+\item Mauritania - Nouakchott
+\item Cape Verde - Praia
+\item Morocco - Rabat
+\item Algeria - Algiers
+\item Tunisia - Tunis
+\item Libya - Tripoli
+\item Egypt - Cairo
+\item Sudan - Khartoum
+\item Eritrea - Asmara
+\item Djibouti - Djibouti
+\item Ethiopia - Addis Ababa
+\item Somalia - Mogadishu
+\item Seychelles - Victoria
+\item Mauritius - Port Louis
+\item Madagascar - Antananarivo
+\item Comoros - Moroni
+\item Tanzania - Dar es Salaam
+\item Kenya - Nairobi
+\item Uganda - Kampala
+\item Rwanda - Kigali
+\item Burundi - Bujumbura
+\item Malawi - Lilongwe
+\item Zambia - Lusaka
+\item Zimbabwe - Harare
+\item Mozambique - Maputo
+\item Swaziland - Mbabane
+\item Lesotho - Maseru
+\item South Africa - Pretoria 
+\item Botswana - Gaborone
+\item Namibia - Windhoek
+\item Angola - Luanda
+\item Congo, Democratic Republic of the - Kinshasa
+\item Congo, Republic of the - Brazzaville
+\item Central African Republic - Bangui
+\item Chad - N'Djamena
+\item Cameroon - Yaounde
+\item Gabon - Libreville
+\item Equatorial Guinea - Malabo
+\item Nigeria - Abuja
+\end{enumerate*}
+\end{multicols}
+
+Total Length Of Non-Round-trip Route $\approx 34,096 km$
 
 \section{Runtime}
-We worked hard, and achieved very little.
+
+The two algorithms, simulated annealing and brute force random permutations where both run in numerous instances over the life of the project. 
+In combined total they were run upwards of fifty times a piece.
+
+The version of the algorithm which used simulated annealing on average found the optimal solution in 6 seconds. 
+While the brute force method often took upwards of
 
@@
 
+I believe the final optimal path's are within a reasonable uncertainty of the actual path.
+Given the big-O of the brute force algorithm, $O(n!)$ or $O( 8.065 \times 10^{67} )$, a runtime complexity of this magnitude is obviously out of my league for finding exact values.
+At a maximum I was able to attempt 100,000,000 permutations of the path.
+
+Two area's of possible improvement I can pinpoint are:
+\begin{enumerate}
+\item The language itself. I chose the python programming language for my TSP implementation.
+The language is interpreted and it's runtime speed can at times be 20\% slower than an equivalent C program.
+\item The random.shuffle implementation. Given how python's random.shuffle API is implemented it takes a set
+of values and shuffles them based on a random number between 0,1 generated by the operating system. However given the size of the set
+, 52 individual points, the likely hood that random.shuffle will generated all permutations without excessive doubles is incredibly unlikely.
+This is the cause of a lot of useless computation time, however because of the size of the data set, caching data sets is not plausible let alone if even possible.
+\end{enumerate}
+
+Re-writing the algorithm in C would be a great performance benefit.
+Also having a proper resources to cache routes or a effective random route generator I believe a more accurate optimal route might be found.
+
+It is also worth mentioning that after completing a large part of the project I discovered the LHK\cite{LHK} high performance TSP solver.
+It's written in straight C and requires the dataset to be in a very specific format.
+Given more time to organize the dataset into the proper format, I believe the LHK solver might produce a more accurate route given a much larger number of iterations. 
+\begin{thebibliography}{99}
+\bibitem{python} {\href{ http://python.org }{ http://python.org }}
+\bibitem{wiki} {\href{ http://en.wikipedia.org/wiki/Latitude_and_longitude_of_cities }{ http://en.wikipedia.org/wiki/Latitude\_and\_longitude\_of\_cities }}
+\bibitem{tsp} Montgomery, John {\it Tackling The Travelling Salesman Problem }{\href{ http://www.psychicorigami.com/category/tsp/ }{ http://www.psychicorigami.com/category/tsp/ } }, {\bf 2007}
+\bibitem{google} {\href{ http://code.google.com/apis/maps/documentation/staticmaps/  }{ http://code.google.com/apis/maps/documentation/staticmaps/ }}
+\bibitem{sa}{\href{ http://www.svengato.com/salesman.html }{ http://www.svengato.com/salesman.html }}
+\bibitem{dis}{\href{ http://www.geesblog.com/2009/01/calculating-distance-between-latitude-longitude-pairs-in-python/ }{ http://www.geesblog.com/2009/01/calculating-distance-between-latitude-longitude-pairs-in-python/ }}
+\bibitem{LHK}{\href{ http://www.akira.ruc.dk/~keld/research/LKH/ }{ http://www.akira.ruc.dk/~keld/research/LKH/ }}
+
+\end{thebibliography}
+\section{Program Output}
+\tiny\begin{verbatim}
+burny ~/src/traveling-salesman $ python tsp.py -v -a anneal -n 1000000 --cooling 10:.9991 ../latlon
+2009-11-02 00:43:38,822 INFO using move_operator: <function reversed_sections at 0x745f0>
+2009-11-02 00:43:38,860 INFO new best score: -111551.944723
+2009-11-02 00:43:38,861 INFO anneal started: score=-111551.944723
+2009-11-02 00:43:38,861 INFO current run at 2 runs
+2009-11-02 00:43:38,861 INFO current run at 4 runs
+2009-11-02 00:43:38,861 INFO new best score: -108938.423206
+2009-11-02 00:43:38,862 INFO new best score: -108255.962603
+2009-11-02 00:43:38,862 INFO current run at 8 runs
+2009-11-02 00:43:38,862 INFO new best score: -107174.578311
+2009-11-02 00:43:38,862 INFO new best score: -107174.578311
+2009-11-02 00:43:38,862 INFO new best score: -106843.737327
+2009-11-02 00:43:38,863 INFO current run at 16 runs
+2009-11-02 00:43:38,864 INFO new best score: -105825.989448
+2009-11-02 00:43:38,864 INFO new best score: -105805.630085
+2009-11-02 00:43:38,864 INFO new best score: -105428.930412
+2009-11-02 00:43:38,864 INFO new best score: -103871.898541
+2009-11-02 00:43:38,865 INFO new best score: -99956.329397
+2009-11-02 00:43:38,865 INFO current run at 32 runs
+2009-11-02 00:43:38,865 INFO new best score: -99252.982749
+2009-11-02 00:43:38,865 INFO new best score: -96688.656537
+2009-11-02 00:43:38,866 INFO new best score: -94637.985111
+2009-11-02 00:43:38,866 INFO new best score: -92809.021849
+2009-11-02 00:43:38,866 INFO new best score: -92567.648622
+2009-11-02 00:43:38,867 INFO new best score: -92341.764572
+2009-11-02 00:43:38,867 INFO new best score: -92019.656620
+2009-11-02 00:43:38,867 INFO new best score: -91759.417996
+2009-11-02 00:43:38,867 INFO new best score: -91657.533796
+2009-11-02 00:43:38,867 INFO new best score: -91615.252931
+2009-11-02 00:43:38,868 INFO new best score: -91510.009964
+2009-11-02 00:43:38,868 INFO new best score: -91409.900530
+2009-11-02 00:43:38,868 INFO new best score: -88486.992310
+2009-11-02 00:43:38,869 INFO new best score: -87865.530081
+2009-11-02 00:43:38,869 INFO current run at 64 runs
+2009-11-02 00:43:38,870 INFO new best score: -84271.022595
+2009-11-02 00:43:38,871 INFO new best score: -83806.794411
+2009-11-02 00:43:38,871 INFO new best score: -83541.482139
+2009-11-02 00:43:38,873 INFO new best score: -83230.783927
+2009-11-02 00:43:38,873 INFO new best score: -82737.304308
+2009-11-02 00:43:38,874 INFO new best score: -81484.288842
+2009-11-02 00:43:38,874 INFO new best score: -78113.454255
+2009-11-02 00:43:38,875 INFO new best score: -78076.761685
+2009-11-02 00:43:38,876 INFO current run at 128 runs
+2009-11-02 00:43:38,876 INFO new best score: -77590.182221
+2009-11-02 00:43:38,876 INFO new best score: -77048.097861
+2009-11-02 00:43:38,876 INFO new best score: -75684.545805
+2009-11-02 00:43:38,877 INFO new best score: -75404.373749
+2009-11-02 00:43:38,877 INFO new best score: -75171.746424
+2009-11-02 00:43:38,878 INFO new best score: -74321.762810
+2009-11-02 00:43:38,879 INFO new best score: -73252.833808
+2009-11-02 00:43:38,879 INFO new best score: -73104.514189
+2009-11-02 00:43:38,881 INFO new best score: -72788.614518
+2009-11-02 00:43:38,881 INFO new best score: -72591.656921
+2009-11-02 00:43:38,881 INFO new best score: -70578.358239
+2009-11-02 00:43:38,882 INFO new best score: -68799.506394
+2009-11-02 00:43:38,883 INFO new best score: -68619.887877
+2009-11-02 00:43:38,884 INFO new best score: -67121.606526
+2009-11-02 00:43:38,884 INFO new best score: -66626.493200
+2009-11-02 00:43:38,884 INFO new best score: -64344.816135
+2009-11-02 00:43:38,884 INFO new best score: -64230.900991
+2009-11-02 00:43:38,885 INFO new best score: -64101.038316
+2009-11-02 00:43:38,885 INFO new best score: -63926.594272
+2009-11-02 00:43:38,886 INFO new best score: -63639.151393
+2009-11-02 00:43:38,886 INFO new best score: -63342.571435
+2009-11-02 00:43:38,886 INFO new best score: -62883.123482
+2009-11-02 00:43:38,886 INFO new best score: -60493.251630
+2009-11-02 00:43:38,887 INFO new best score: -58493.381705
+2009-11-02 00:43:38,888 INFO new best score: -58364.904831
+2009-11-02 00:43:38,888 INFO new best score: -57861.729512
+2009-11-02 00:43:38,889 INFO new best score: -57785.659587
+2009-11-02 00:43:38,889 INFO new best score: -56237.272807
+2009-11-02 00:43:38,889 INFO current run at 256 runs
+2009-11-02 00:43:38,890 INFO new best score: -52752.018487
+2009-11-02 00:43:38,890 INFO new best score: -52058.361255
+2009-11-02 00:43:38,890 INFO new best score: -51827.173423
+2009-11-02 00:43:38,891 INFO new best score: -50650.559834
+2009-11-02 00:43:38,894 INFO new best score: -50540.078096
+2009-11-02 00:43:38,895 INFO new best score: -50317.193191
+2009-11-02 00:43:38,895 INFO new best score: -48975.975547
+2009-11-02 00:43:38,896 INFO new best score: -48849.972600
+2009-11-02 00:43:38,898 INFO new best score: -48838.291771
+2009-11-02 00:43:38,902 INFO new best score: -48368.567964
+2009-11-02 00:43:38,903 INFO new best score: -46559.324605
+2009-11-02 00:43:38,904 INFO new best score: -46558.360448
+2009-11-02 00:43:38,905 INFO new best score: -46522.611077
+2009-11-02 00:43:38,905 INFO new best score: -46422.331395
+2009-11-02 00:43:38,906 INFO new best score: -44981.634651
+2009-11-02 00:43:38,907 INFO new best score: -44820.384544
+2009-11-02 00:43:38,907 INFO new best score: -44643.673209
+2009-11-02 00:43:38,909 INFO new best score: -44004.657313
+2009-11-02 00:43:38,909 INFO new best score: -43563.669128
+2009-11-02 00:43:38,910 INFO new best score: -43510.757103
+2009-11-02 00:43:38,912 INFO current run at 512 runs
+2009-11-02 00:43:38,916 INFO new best score: -42977.735342
+2009-11-02 00:43:38,916 INFO new best score: -41456.278999
+2009-11-02 00:43:38,917 INFO new best score: -41383.345365
+2009-11-02 00:43:38,918 INFO new best score: -40087.086741
+2009-11-02 00:43:38,919 INFO new best score: -39434.670603
+2009-11-02 00:43:38,919 INFO new best score: -39219.990542
+2009-11-02 00:43:38,921 INFO new best score: -38293.656546
+2009-11-02 00:43:38,924 INFO new best score: -36738.445888
+2009-11-02 00:43:38,926 INFO new best score: -36649.112566
+2009-11-02 00:43:38,932 INFO new best score: -36591.647201
+2009-11-02 00:43:38,933 INFO new best score: -36238.421358
+2009-11-02 00:43:38,935 INFO new best score: -36196.494534
+2009-11-02 00:43:38,938 INFO new best score: -36014.634524
+2009-11-02 00:43:38,940 INFO new best score: -35991.037879
+2009-11-02 00:43:38,941 INFO new best score: -35889.038030
+2009-11-02 00:43:38,946 INFO new best score: -35667.660601
+2009-11-02 00:43:38,953 INFO new best score: -35491.868203
+2009-11-02 00:43:38,955 INFO current run at 1024 runs
+2009-11-02 00:43:38,958 INFO new best score: -35322.422385
+2009-11-02 00:43:38,971 INFO new best score: -35296.389295
+2009-11-02 00:43:38,972 INFO new best score: -34799.198370
+2009-11-02 00:43:38,974 INFO new best score: -34655.194786
+2009-11-02 00:43:38,977 INFO new best score: -34608.069227
+2009-11-02 00:43:38,979 INFO new best score: -34539.947513
+2009-11-02 00:43:38,979 INFO new best score: -33247.031299
+2009-11-02 00:43:38,979 INFO new best score: -32367.878943
+2009-11-02 00:43:38,987 INFO new best score: -31805.170969
+2009-11-02 00:43:38,996 INFO new best score: -31465.390351
+2009-11-02 00:43:38,996 INFO new best score: -31399.347724
+2009-11-02 00:43:38,999 INFO new best score: -30749.273747
+2009-11-02 00:43:39,003 INFO new best score: -30425.747441
+2009-11-02 00:43:39,005 INFO new best score: -29257.455498
+2009-11-02 00:43:39,006 INFO new best score: -29215.783892
+2009-11-02 00:43:39,008 INFO new best score: -28975.291472
+2009-11-02 00:43:39,012 INFO new best score: -28192.624760
+2009-11-02 00:43:39,018 INFO new best score: -27934.421265
+2009-11-02 00:43:39,026 INFO new best score: -27688.247659
+2009-11-02 00:43:39,035 INFO current run at 2048 runs
+2009-11-02 00:43:39,037 INFO new best score: -27390.644387
+2009-11-02 00:43:39,057 INFO new best score: -26793.922438
+2009-11-02 00:43:39,057 INFO new best score: -26696.244150
+2009-11-02 00:43:39,061 INFO new best score: -26404.136400
+2009-11-02 00:43:39,083 INFO new best score: -26404.136400
+2009-11-02 00:43:39,094 INFO new best score: -25926.038197
+2009-11-02 00:43:39,108 INFO new best score: -25368.028231
+2009-11-02 00:43:39,115 INFO new best score: -25312.586966
+2009-11-02 00:43:39,132 INFO new best score: -25251.911826
+2009-11-02 00:43:39,137 INFO new best score: -25202.999277
+2009-11-02 00:43:39,145 INFO new best score: -24692.239918
+2009-11-02 00:43:39,158 INFO new best score: -24692.239918
+2009-11-02 00:43:39,187 INFO new best score: -24206.877575
+2009-11-02 00:43:39,195 INFO current run at 4096 runs
+2009-11-02 00:43:39,223 INFO new best score: -24159.595396
+2009-11-02 00:43:39,231 INFO new best score: -24088.719891
+2009-11-02 00:43:39,245 INFO new best score: -23972.734353
+2009-11-02 00:43:39,260 INFO new best score: -23906.426074
+2009-11-02 00:43:39,289 INFO new best score: -23901.325907
+2009-11-02 00:43:39,305 INFO new best score: -23286.130255
+2009-11-02 00:43:39,309 INFO new best score: -23076.669258
+2009-11-02 00:43:39,310 INFO new best score: -22945.476729
+2009-11-02 00:43:39,320 INFO new best score: -22943.882305
+2009-11-02 00:43:39,370 INFO new best score: -22097.219253
+2009-11-02 00:43:39,378 INFO new best score: -21951.069041
+2009-11-02 00:43:39,512 INFO current run at 8192 runs
+2009-11-02 00:43:39,769 INFO new best score: -21949.474618
+2009-11-02 00:43:40,134 INFO current run at 16384 runs
+2009-11-02 00:43:40,367 INFO new best score: -21819.103878
+2009-11-02 00:43:40,368 INFO new best score: -21814.003710
+2009-11-02 00:43:40,430 INFO new best score: -21812.409286
+2009-11-02 00:43:40,442 INFO new best score: -21812.409286
+2009-11-02 00:43:41,231 INFO new best score: -21729.582347
+2009-11-02 00:43:41,262 INFO new best score: -21604.015041
+2009-11-02 00:43:42,573 INFO new best score: -21549.955257
+2009-11-02 00:43:43,565 INFO current run at 32768 runs
+2009-11-02 00:43:43,652 INFO new best score: -21549.955257
+2009-11-02 00:43:45,122 INFO new best score: -21549.955257
+2009-11-02 00:43:48,326 INFO current run at 65536 runs
+2009-11-02 00:43:54,024 INFO current run at 131072 runs
+2009-11-02 00:44:05,620 INFO current run at 262144 runs
+2009-11-02 00:44:29,461 INFO current run at 524288 runs
+2009-11-02 00:45:13,352 INFO final temperature: 1.232621
+2009-11-02 00:45:13,352 INFO anneal finished: num_evaluations=1000000, best_score=-21549.955257
+Benin - Porto-Novo
+Togo - Lome
+Ghana - Accra
+Niger - Niamey
+Burkina Faso - Ouagadougou
+Cote d'Ivoire - Yamoussoukro
+Liberia - Monrovia
+Sierra Leone - Freetown
+Guinea - Conakry
+Guinea-Bissau - Bissau
+The Gambia - Banjul
+Senegal - Dakar
+Mauritania - Nouakchott
+Cape Verde - Praia
+Morocco - Rabat
+Algeria - Algiers
+Tunisia - Tunis
+Libya - Tripoli
+Egypt - Cairo
+Sudan - Khartoum
+Eritrea - Asmara
+Djibouti - Djibouti
+Ethiopia - Addis Ababa
+Somalia - Mogadishu
+Seychelles - Victoria
+Mauritius - Port Louis
+Madagascar - Antananarivo
+Comoros - Moroni
+Tanzania - Dar es Salaam
+Kenya - Nairobi
+Uganda - Kampala
+Rwanda - Kigali
+Burundi - Bujumbura
+Malawi - Lilongwe
+Zambia - Lusaka
+Zimbabwe - Harare
+Mozambique - Maputo
+Swaziland - Mbabane
+Lesotho - Maseru
+South Africa - Pretoria 
+Botswana - Gaborone
+Namibia - Windhoek
+Angola - Luanda
+Congo, Democratic Republic of the - Kinshasa
+Congo, Republic of the - Brazzaville
+Central African Republic - Bangui
+Chad - N'Djamena
+Cameroon - Yaounde
+Gabon - Libreville
+Equatorial Guinea - Malabo
+Nigeria - Abuja
+21186.3444977 mi
+\end{verbatim}
+
 \end{document}

### 2009-11-02 anealing/sa.py

@@ -27,7 +27,7 @@ class ObjectiveFunction:
             logging.info('new best score: %f',self.best_score)
             if score > (-341):
               getimg(self.coords,self.best)
-              os.popen("eog path.png")
+              os.popen("open path.png")
 
         return score
 
@@ -83,7 +83,7 @@ def anneal(init_function,move_operator,objective_function,max_evaluations,start_
     return (num_evaluations,best_score,best)
 
 def getimg(coords, best):
-    url = "http://maps.google.com/maps/api/staticmap?path=color:orange|weight:4"
+    url = "http://maps.google.com/maps/api/staticmap?path=color:0x000000ff|weight:4"
 
     for i in best:
         url = url + "|" + str(coords[i][0]) + "," + str(coords[i][1])
@@ -96,6 +96,6 @@ def getimg(coords, best):
             url = url + "|" + str(coords[i][0]) + "," + str(coords[i][1])
 
     url = url + "&size=1024x1024&sensor=false&key=ABQIAAAAzr2EBOXUKnm_jVnk0OJI7xSsTL4WIgxhMZ0ZK_kHjwHeQuOD4xQJpBVbSrqNn69S6DOTv203MQ5ufA"
-    urlretrieve(url,"points.png")
+    #urlretrieve(url,"points.png")
 
 

### 2009-11-02 test/tsp.py

@@ -71,7 +71,7 @@ def read_coords(coord_file):
     return coords
 
 def order(best):
-    file = open('cities','r')
+    file = open('../cities','r')
     cities = []
     for city in file:
       cities.append(city)
@@ -237,6 +237,7 @@ def main():
     iterations,score,best=run_algorithm(init_function,move_operator,objective_function,max_iterations,coords)
     # output results
     #print order(best)
+    print order(best)
     print iterations,score,best
     print str(calculatePathLen(best,coords,tsp)) + " mi"
     

### 2009-11-02 paper/paper.tex
paper

@@
 the shortest round trip path a traveling salesperson can take
-to visit each vertex of a given graph.  They are usually
-implemented using a genetic algorithm. Our salesperson
+to visit each vertex of a given graph. Our salesperson
 happens to be traveling to the capitals of every country in Africa that is a recognized member of the United Nations.
@@
 
+\newpage      
 \section{The Problem}
+\bf{G-3(b)****. }\normalfont A Traveling Salesman Problem is the task of finding 
+the shortest roundtrip path a traveling salesperson can take 
+to visit each vertex of a given graph.  They are usually 
+implemented using a genetic algorithm.
 
-  Attempt to find the shortest i.e. cheapest complete path around all African Nations so our sales person can travel the most efficient route possible.
+\begin{itemize}
+\item You may choose only one of the following sets of vertices. \newline
+\small
+\bf(a) \normalfont \small S = \{capitals of the 50 states in the US \} \newline
+\bf(b) \normalfont \small S = \{capitals of every country in Africa that is a recognized member of the United Nations\} \newline
+\bf(c) \normalfont \small S = \{capitals of every country in Asia that is a recognized member of the United Nations\} \newline
+\bf(d) \normalfont \small S = \{capitals of every country in Europe that is a recognized member of the United Nations\}
+\normalfont
+\item Find the shortest roundtrip path that visits all the vertices in S, traveling along geodesics (that is, 
+straight lines over the surface of the earth).  [Important:  Do not add additional cities to your set.] 
+\item Include the optimal distance (rounded to the nearest kilometer) and a map of the optimal route.   
+\item Provide a list of countries (in alphabetical order) and capitals in your set. 
+\item Find the shortest path if we drop the condition that the path must be a roundtrip. 
+\end{itemize}
 
@@
 
-The version of the algorithm which used simulated annealing on average found the optimal solution in 6 seconds. 
-While the brute force method often took upwards of
+The version of the algorithm which used simulated annealing on average found the optimal solution in approximately 6 seconds. 
+While the brute force algorithm delivered the same result or close to it in upwards of 2 to 10 minutes.
+Given the internal implementation of the algorithm, it's performance is directly linked with the first random number used to seed the initial route.
+Multiple runs of the brute force algorithm another one another often result in drastically different run times to find the same solution.
+This was beyond my control on the implementation side of things.
 
